---
title: "Second-hand cars price EDA & prediction"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(mice)
library(ggcorrplot)
library(ggthemes)
library(scales)


theme_set(theme_light())

```

```{r}
data_raw <- read.csv("carsData24.csv")
View(data_raw)

data <- subset(data_raw, select = -X)
head(data, 10)
```
```{r}
md.pattern(data)
```
```{r}
data <- na.omit(data)
colSums(is.na(data))

md.pattern(data)
```

# Exploration

## Correlation
```{r}
data.numeric <- data %>% 
  dplyr::select(where(is.numeric))

cor.data <- cor(data.numeric)

ggcorrplot(cor.data, lab = TRUE)
```

## Response Variable : price
```{r}

data %>% 
  ggplot() + 
  geom_histogram(aes(price, ..density..), bins=50) +
  geom_density(aes(price)) +
  scale_x_continuous(labels = label_dollar())

```
## year
```{r}
data %>% ggplot(aes(year, price)) +
  geom_point() +
  geom_smooth(method = "loess") +
  scale_y_continuous(labels = label_dollar())
```
```{r}
data %>% 
  filter(year > 1980)  %>% 
  ggplot(aes(year, price)) +
  geom_point() +
  geom_smooth(method = "loess") +
  scale_y_continuous(labels = label_dollar())
```

## Cylinders
```{r}
data %>% ggplot(aes(cylinders, price)) +
  geom_col()  +
  scale_y_continuous(labels = label_dollar())
```
## Odometer
```{r}
data %>% ggplot(aes(odometer, price)) +
  geom_point() +
  geom_smooth(method = "loess") +
  scale_y_continuous(labels = label_dollar())
```
Interseting that prices stabilise after odometer readings of 150+. Expectation is that it should keep on decreasing. 

```{r}
data %>% filter(odometer>150)
```
One explanation is that the conditions, on account of them being "good" or better, plays a part. Let's try to see if that's the case indeed.
```{r}
data %>% 
  ggplot(aes(x=odometer, y=price, group = condition, color = condition)) +
  geom_point(stat = "summary", fun.y = "mean") +
  scale_y_continuous(labels = label_dollar())
```

## State
```{r}
data %>% 
  mutate(state = fct_reorder(state, price)) %>% 
           ggplot(aes(state, price)) + 
           geom_boxplot() +
  scale_y_continuous(labels = label_dollar()) +
  coord_flip() 
```

## Type of Car
```{r}
data %>% 
  mutate(type = fct_reorder(type, price)) %>% 
  ggplot(aes(type, price)) +
  geom_boxplot() +
  scale_y_continuous(labels = label_dollar())
```
## Size
```{r}
data %>% 
  mutate(size = fct_reorder(size, price)) %>% 
  ggplot(aes(size, price)) +
  geom_boxplot() +
  scale_y_continuous(labels = label_dollar())
```
For some reason, sub-compact has the highest median price. Let's see what cars are of that size category. 
```{r}
data %>% 
  filter(size == 'sub-compact')
```
Interesting, there is only one car.

## Fuel
```{r}
data %>% ggplot(aes(fuel, price)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.1) +
  labs(x = "Fuel Type",
       y = "Price") +
  scale_y_continuous(labels = label_dollar())
```
## Year & Condition
```{r}
data %>% 
  ggplot(aes(year, price, color = condition)) +
  geom_point() +
  geom_text(aes(label = manufacturer), vjust = 1, hjust = 1,check_overlap = TRUE) +
  scale_y_continuous(labels = label_dollar())
```
Looks like a few old cars are in 'excellent' condition. Interestingly, Chevrolet is the manufacturer for two out of three of those cars in that condition.

## Transmission
```{r}
data %>% 
  ggplot(aes(transmission, price)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.1) +
  scale_y_continuous(labels = label_dollar())
```
```{r}
data %>% 
  group_by(transmission) %>% 
  summarize(mean_price = mean(price),
            median_price = median(price))
```

## Manufacturer
```{r}
data %>% 
  mutate(manufacturer = fct_reorder(manufacturer, price)) %>% 
  ggplot(aes(manufacturer, price)) +
  geom_boxplot() +
  scale_y_continuous(labels = label_dollar()) +
  coord_flip() +
  labs(x = "Manufacturer",
       y = "Price ($)",
       title = "Prices of cars vs the Manufacturing companies")
```
## Paint colour
```{r}
data %>% 
  mutate(paint_color = fct_reorder(paint_color, price)) %>% 
  ggplot(aes(paint_color, price)) +
  geom_boxplot() +
  scale_y_continuous(labels = label_dollar()) +
  coord_flip()
```
```{r}
data %>%  
  group_by(manufacturer, paint_color) %>% 
  summarize(mean_price = mean(price),
            median_price = median(price)) %>% 
  arrange(desc(median_price))
```
# Fitting models

## Numerical Regression Model
```{r}
#data %>% 
#  mutate(paint_color = fct_lump(paint_color, 5),
#         paint_color = fct_reorder(paint_color, price)) %>% 
#  ggplot(aes(paint_color, price)) +
#  geom_boxplot() +
#  scale_y_continuous(labels = label_dollar()) +
#  coord_flip()
```

```{r}
lm.num <- lm(price ~ year + cylinders + odometer, data)
summary(lm.num)
```
```{r}
plot(fitted(lm.num), rstandard(lm.num),
     xlab = "Fitted", ylab = "Standardised Residuals")

lines(lowess(fitted(lm.num), rstandard(lm.num)), col = "red")

abline(h=qnorm(c(0.025,0.975),0,sd(rstandard(lm.num))),
       col="grey", lty=2)
```
```{r}
# QQ plots
qqnorm(residuals(lm.num))
qqline(residuals(lm.num))
```
## Dummy variables Regression model
```{r}
## Creating dummy variables
data.reg <- data
data.reg$manufacturer <- as.factor(data.reg$manufacturer)
data.reg$model <- as.factor(data.reg$model)
data.reg$condition <- as.factor(data.reg$condition)
data.reg$fuel <- as.factor(data.reg$fuel)
data.reg$transmission <- as.factor(data.reg$transmission)
data.reg$size <- as.factor(data.reg$size) 
data.reg$type <- as.factor(data.reg$type)
data.reg$paint_color <- as.factor(data.reg$paint_color)
data.reg$state <- as.factor(data.reg$state)
```

```{r}
# Excluding model
lm.char <- lm(price ~ manufacturer + condition + fuel + transmission + size + type + paint_color + state, data = data)
summary(lm.char)
```
```{r}
plot(lm.char)
```

